A Bidirected String Graph Model of Genome Assembly

### 2012-11-29 presentation/presentation.tex
f.E g.E slide

\begin{frame}{Overlaps between reads}
	Consider two reads, $f$ and $g$, and their reverse-complements $f'$ and $g'$.
	The reads $f$ and $g$ are said to {\bf exactly overlap} by some length
	$\LengthVar$ iff at least one of the following is true:
	\begin{itemize}
		\item The last $\LengthVar$ bp of $f$ exactly match the first
		$\LengthVar$ bp of $g$.
		\item The last $\LengthVar$ bp of $f$ exactly match the first
		$\LengthVar$ bp of $g'$.
		\item The last $\LengthVar$ bp of $f'$ exactly match the first
		$\LengthVar$ bp of $g$.
		\item The last $\LengthVar$ bp of $f'$ exactly match the first
		$\LengthVar$ bp of $g'$.
	\end{itemize}
	An exact overlap $o$ is fully described by the tuple $(f, g, f_{begin},
	g_{begin}, L, rc)$.
	%there $f$ and $g$ are the reads of the overlap,
	%$f_{beg}$ and $g_{beg}$ and the starting position of the overlap on each
	%read, $len$ is the length of the overlap, and $rc$ indicates whether the
	%overlap is reverse-complement or not.
\end{frame}

@@

\begin{frame}{Build the fragment string assembly graph (cont.)}
	\begin{itemize}
		\item Consider three reads $f$, $g$, and $h$, that overlap as follows:
	\end{itemize}
	\vspace{-0.4cm}
	\begin{center}
		\begin{tikzpicture}[>=\ReadArrowType]
			\draw[->,style=very thick] (0,   0.6) node[anchor=east] {f} -- (3.0, 0.6);
			\draw[->,style=very thick] (1.0, 0.3) node[anchor=east] {g} -- (4.0, 0.3);
			\draw[<-,style=very thick] (2.0, 0.0) node[anchor=east] {h} -- (5.0, 0.0);
		\end{tikzpicture}
	\end{center}
	\vspace{-0.5cm}
	\begin{itemize}
		\item {There are 3 overlaps so there will be 3 edges, and we'd like for
		the graph to allow exactly the following four walks:}
		\vspace{-0.3cm}
		\begin{multicols}{2}
			\begin{itemize}
				\item $f \to g \to h$
				\item $h \to g \to f$
				\item $f \to h$
				\item $h \to f$
			\end{itemize}
		\end{multicols}
	\end{itemize}
	\begin{tabular}{ccp{3.3cm}}
		{\small Undirected graph?}
			&
			\begin{tikzpicture}[>=triangle 45]
				\tikzstyle{every node} = [circle,fill=blue!20,draw=black];
				\node (f) at (0, 0) {f};
				\node (g) at (1.7, 0) {g};
				\node (h) at (3.4, 0) {h};
				\draw[-,style=very thick] (f) edge (g);
				\draw[-,style=very thick] (g) edge (h);
				\draw[-,style=very thick,bend left] (f) edge (h);
			\end{tikzpicture}
			& 
			\vspace{-0.8cm}
			{\footnotesize No; nothing indicates we can't walk \mbox{$g \to h \to
			f$}.}
			\\
		{\small Directed graph?}
			&
			\begin{tikzpicture}[>=triangle 45]
				\tikzstyle{every node} = [circle,fill=blue!20,draw=black];
				\node (f) at (0, 0) {f};
				\node (g) at (1.7, 0) {g};
				\node (h) at (3.4, 0) {h};
				\draw[->,style=very thick] (f) edge (g);
				\draw[->,style=very thick] (g) edge (h);
				\draw[->,style=very thick,bend left] (f) edge (h);
			\end{tikzpicture}
			&
			\vspace{-0.8cm}
			{\footnotesize
				No; this doesn't allow the walks $h \to g \to f$ and $h \to f$.}
			\\
	\end{tabular}
\end{frame}

@@

%\begin{frame}{Walking through the bidirected string graph (cont.)}
	%\begin{minipage}{0.5\textwidth}
		%\begin{tikz
	%\end{minipage}
%\end{frame}

>>>>

\begin{frame}{Overlaps between reads}
	Consider two reads, $f$ and $g$, and their reverse-complements $f'$ and $g'$.
	The reads $f$ and $g$ are said to {\bf exactly overlap} by some length
	$\LengthVar$ iff at least one of the following is true:
	\begin{itemize}
		\item The last $\LengthVar$ bp of $f$ exactly match the first
		$\LengthVar$ bp of $g$.
		\item The last $\LengthVar$ bp of $f$ exactly match the first
		$\LengthVar$ bp of $g'$.
		\item The last $\LengthVar$ bp of $f'$ exactly match the first
		$\LengthVar$ bp of $g$.
		\item The last $\LengthVar$ bp of $f'$ exactly match the first
		$\LengthVar$ bp of $g'$.
	\end{itemize}
	An overlap $o$ is fully described by the 7-tuple $(f, g, f_{begin}, f_{end},
	g_{begin}, g_{end}, rc)$.
	%there $f$ and $g$ are the reads of the overlap,
	%$f_{beg}$ and $g_{beg}$ and the starting position of the overlap on each
	%read, $len$ is the length of the overlap, and $rc$ indicates whether the
	%overlap is reverse-complement or not.
\end{frame}

@@

\begin{frame}{Build the fragment string assembly graph (cont.)}
	\begin{itemize}
		\item Consider three reads $f$, $g$, and $h$, that overlap as follows:
	\end{itemize}
	\vspace{-0.4cm}
	\begin{center}
		\begin{tikzpicture}[>=\ReadArrowType]
			\draw[->,style=very thick] (0,   0.6) node[anchor=east] {f} -- (3.0, 0.6);
			\draw[->,style=very thick] (1.0, 0.3) node[anchor=east] {g} -- (4.0, 0.3);
			\draw[<-,style=very thick] (2.0, 0.0) node[anchor=east] {h} -- (5.0, 0.0);
		\end{tikzpicture}
	\end{center}
	\vspace{-0.5cm}
	\begin{itemize}
		\item {There are 3 overlaps so there will be 3 edges, and we'd like for
		the graph to allow exactly the following four walks:}
		\vspace{-0.3cm}
		\begin{multicols}{2}
			\begin{itemize}
				\item $f \to g \to h$
				\item $h \to g \to f$
				\item $f \to h$
				\item $h \to f$
			\end{itemize}
		\end{multicols}
	\end{itemize}
	\begin{tabular}{ccp{3.4cm}}
		{\small Undirected graph?}
			&
			\begin{tikzpicture}[>=triangle 45]
				\tikzstyle{every node} = [circle,fill=blue!20,draw=black];
				\node (f) at (0, 0) {f};
				\node (g) at (1.7, 0) {g};
				\node (h) at (3.4, 0) {h};
				\draw[-,style=very thick] (f) edge (g);
				\draw[-,style=very thick] (g) edge (h);
				\draw[-,style=very thick,bend left] (f) edge (h);
			\end{tikzpicture}
			& 
			\vspace{-1.1cm}
			{\footnotesize Not what we want; nothing indicates we can't walk \mbox{$g \to h \to
			f$}.}
			\\
		{\small Directed graph?}
			&
			\begin{tikzpicture}[>=triangle 45]
				\tikzstyle{every node} = [circle,fill=blue!20,draw=black];
				\node (f) at (0, 0) {f};
				\node (g) at (1.7, 0) {g};
				\node (h) at (3.4, 0) {h};
				\draw[->,style=very thick] (f) edge (g);
				\draw[->,style=very thick] (g) edge (h);
				\draw[->,style=very thick,bend left] (f) edge (h);
			\end{tikzpicture}
			&
			\vspace{-0.8cm}
			{\footnotesize
				Not what we want; this doesn't allow the walks $h \to g \to f$ and $h \to f$.}
			\\
	\end{tabular}
\end{frame}

@@

\begin{frame}{Walking through the bidirected string graph (cont.)}
	\vspace{-0.7cm}
	\begin{center}
		\begin{tabular}{p{3.4cm}c}
			\vspace{-0.4cm}
			{\small a.) A bidirected string graph fragment} &
			\begin{tikzpicture}[>=triangle 45,scale=1.5]
				\tikzstyle{every node} = [circle,fill=blue!20,draw=black];
				\node (f) at (0, 0) {f};
				\node (g) at (1.7, 0) {g};
				\draw[>->,style=very thick] (f) edge (g);
			\end{tikzpicture}
			\\
			\vspace{-1.5cm}
			{\small b.) The corresponding overlap (labeled)} &
			\begin{tikzpicture}[>=\ReadArrowType,scale=0.8]
				\draw[->,very thick] (0, 0) node[anchor=east] {$f.B$}
						-> (4, 0) node[anchor=west] {$f.E$};
				\draw[->,very thick] (2, -0.7) node[anchor=east] {$g.B$}
						-> (6, -0.7) node[anchor=west] {$g.E$};
				\draw (2, -1.05) node[anchor=north] {$o.g_{begin}$}
						-- (2, 0.35) node[anchor=south] {$o.f_{begin}$};
				\draw (4, -1.05) node[anchor=north] {$o.g_{end}$}
						-- (4, 0.35) node[anchor=south] {$o.f_{end}$};
			\end{tikzpicture}
		\end{tabular}
	\end{center}
	\begin{tabular}{c|c}
		walk & interpretation \\ \hline
		$f \to g$ & $f.E \to g.E$ with sequence $g[o.g_{end} + 1 \dots ${\tt
		length}$(g)]$ \\
		$g \to f$ & $g.B \to f.B$ with sequence $f[o.g_{begin} - 1 \dots 0]$ \\
	\end{tabular}
	\vspace{0.5cm}
	\begin{itemize}
		\item If we take {\bf each end} of {\bf each read} as a vertex, we can
		make a directed graph equivalent to the bidirected graph.
	\end{itemize}
\end{frame}
